```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(janitor)
library(ggplot2)
library(dplyr)
library(grid)
```

```{r}
#| label: load-data
college_career_data <- read_csv("datasets/college_career.csv")
garden_data <- read_csv("datasets/garden_revised.csv")
library_data <- read_csv("datasets/library_rev.csv")
fair_data <- read_csv("datasets/state_fair.csv")
credit_data <- read_csv("datasets/credit.csv")
school_supplies_data <- read_csv("datasets/school_revised.csv")
```
```{r}
#| label: data-processing

garden_data <- garden_data |>
  rename(site = site_location) |>
  mutate(site = case_when(site == "Northside" ~ "North Side",
                          site == "Westside" ~ "West Side",
                          site == "Centralish" ~ "Central City",
                          site == "Southside" ~ "South Side",
                          site == "Eastside" ~ "East Side")) |>
  mutate(produce_handled_per_volunteer = if_else(site == "West Side", 
                                                 2.5, 
                                                 produce_handled_per_volunteer))

college_career_data <- college_career_data |>
  mutate(class_year = case_when(class_year == "Freshmen" ~ "Freshman",
                                class_year == "Sophomores" ~ "Sophomore",
                                class_year == "Juniors" ~ "Junior",
                                class_year == "Seniors" ~ "Senior")) |>
  mutate(satisfaction_career_center = if_else(class_year == "Freshman",
                                              4, satisfaction_career_center))

fair_data <- fair_data |>
  mutate(pct_attending_all_days = case_match(visitor_type, 
                                             "Student" ~ 60, 
                                             "VIP" ~ 80,
                                             .default = pct_attending_all_days))

credit_data <- credit_data |>
  clean_names()

school_supplies_data <- school_supplies_data |>
  mutate(notebook_type = ifelse(notebook_type == "Blank", 
                                "Layflat", notebook_type))
```


## FUNCTIONS

#| label: lollipop-chart
# TEMPLATE FUNCTION: LOLLIPOP CHART
```{r}
create_lollipop_chart <- function(data, 
                                   category_col, 
                                   value_col,
                                   title,
                                   y_label,
                                   category_order = NULL,
                                  ymin = NULL,
                                  ymax = NULL,
                                   y_break_interval = 50,
                                   color = "chocolate3",
                                   output_file = NULL,
                                   x_text = 14,
                                   y_text = 14,
                                   title_size = 16,
                                  y_axis_labels = NULL) {
  
  # Prepare the data
  plot_data <- data |>
    select(category = {{category_col}}, value = {{value_col}})
  
  # Set category order if provided
  if (!is.null(category_order)) {
    plot_data <- plot_data |>
      mutate(category = factor(category, levels = category_order))
  }
  
  # Calculate y-axis upper limit
  if (is.null(ymax)) {
    ymax <- max(plot_data$value, na.rm = TRUE)
    if (!is.null(y_break_interval)) {
      yupper <- ceiling(ymax / y_break_interval) * y_break_interval
    } else {
      yupper <- ymax
    }
  } else {
    yupper <- ymax  # Use the provided max
  }
  
  # Create the plot
  p <- ggplot(plot_data, aes(x = category, y = value)) +
    geom_segment(aes(xend = category, yend = 0),
                 color = color, linewidth = 2.3) +
    geom_point(color = color, size = 9) +
    labs(
      title = title,
      x = NULL,
      y = NULL
    ) +
    scale_y_continuous(
      limits = c(0, yupper),
      #breaks = seq(0, yupper, by = y_break_interval),
      labels = if (!is.null(y_axis_labels)) y_axis_labels else waiver(),
      expand = expansion(mult = c(0, 0.05))
    ) +
    theme_classic() +
    theme(
      plot.title = element_text(
        size = title_size,
        face = "bold", 
        vjust = 1,
        hjust = 0.5, 
        margin = margin(t = -20, b = 60)
      ),
      plot.margin = margin(45, 40, 25, 40),
      axis.text.x = element_text(size = x_text, color = "black"),
      axis.text.y = element_text(size = y_text, color = "black"),
      axis.ticks.x = element_blank(),
      axis.ticks.y = element_line(color = "black"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    ) +
    #PROBLEMATIC AREA OF Y-AXIS CENTERING
    annotation_custom(
      grob = textGrob(
        label = y_label, 
        hjust = 0.5,
        vjust = -0.9,
        gp = gpar(fontsize = 14)
      ),
      ymin = yupper,
      ymax = yupper,
      xmin = 0.5,
      xmax = 0.5
    ) +
    coord_cartesian(clip = "off")
  
  # Build the plot and turn off clipping
  gt <- ggplot_gtable(ggplot_build(p))
  gt$layout$clip[gt$layout$name == "panel"] <- "off"
  
  # Save if output file specified
  if (!is.null(output_file)) {
    ggsave(output_file, plot = gt, device = "png", width = 8, height = 5)
  }
  
  # Draw and return
  grid.draw(gt)
  invisible(gt)
}

```

```{r}
#| label: bar-chart
# TEMPLATE FUNCTION: BAR CHART
create_bar_chart <- function(data, 
                              category_col, 
                              value_col,
                              title,
                              y_label,
                              category_order = NULL,
                             ymin = NULL,
                             ymax = NULL,
                              y_break_interval = 50,
                              color = "chocolate3",
                              output_file = NULL,
                              x_text = 14,
                              y_text = 14,
                              title_size = 16,
                             y_axis_labels = NULL) {
  
  # Prepare the data
  plot_data <- data |>
    select(category = {{category_col}}, value = {{value_col}})
  
  # Set category order if provided
  if (!is.null(category_order)) {
    plot_data <- plot_data |>
      mutate(category = factor(category, levels = category_order))
  }
  
  # Calculate y-axis upper limit
  if (is.null(ymax)) {
    ymax <- max(plot_data$value, na.rm = TRUE)
    if (!is.null(y_break_interval)) {
      yupper <- ceiling(ymax / y_break_interval) * y_break_interval
    } else {
      yupper <- ymax
    }
  } else {
    yupper <- ymax  # Use the provided max
  }
  
  # Create the plot
  p <- ggplot(plot_data, aes(x = category, y = value)) +
    geom_col(fill = color, width = 0.4) +
    labs(
      title = title,
      x = NULL,
      y = NULL
    ) +
    scale_y_continuous(
      limits = c(0, yupper),
      breaks = seq(0, yupper, by = y_break_interval),
      labels = if (!is.null(y_axis_labels)) y_axis_labels else waiver(),
      expand = expansion(mult = c(0, 0.05))
    ) +
    theme_classic() +
    theme(
      plot.title = element_text(
        size = title_size,
        face = "bold", 
        vjust = 1,
        hjust = 0.5, 
        margin = margin(t = -20, b = 60)
      ),
      plot.margin = margin(45, 40, 25, 40),
      axis.text.x = element_text(size = x_text, color = "black"),
      axis.text.y = element_text(size = y_text, color = "black"),
      axis.ticks.x = element_blank(),
      axis.ticks.y = element_line(color = "black"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    ) +
    annotation_custom(
      grob = textGrob(
        label = y_label, 
        hjust = 0.5,
        vjust = -0.9,
        gp = gpar(fontsize = 14)
      ),
      ymin = yupper,
      ymax = yupper,
      xmin = 0.5,
      xmax = 0.5
    ) +
    coord_cartesian(clip = "off")
  
  # Build the plot and turn off clipping
  gt <- ggplot_gtable(ggplot_build(p))
  gt$layout$clip[gt$layout$name == "panel"] <- "off"
  
  # Save if output file specified
  if (!is.null(output_file)) {
    ggsave(output_file, plot = gt, device = "png", width = 8, height = 5)
  }
  
  # Draw and return
  grid.draw(gt)
  invisible(gt)
}
```

```{r}

#| label: dot-plot
# TEMPLATE FUNCTION: DOT PLOT
create_dot_plot <- function(data, 
                             category_col, 
                             value_col,
                             title,
                             y_label,
                             category_order = NULL,
                            ymin = NULL,
                            ymax = NULL,
                             y_break_interval = 0.5,
                             color = "chocolate3",
                             output_file = NULL,
                             x_text = 14,
                             y_text = 14,
                             title_size = 16,
                            y_axis_labels = NULL) {
  
  # Prepare the data
  plot_data <- data |>
    select(category = {{category_col}}, value = {{value_col}})
  
  # Set category order if provided
  if (!is.null(category_order)) {
    plot_data <- plot_data |>
      mutate(category = factor(category, levels = category_order))
  }
  
  # Calculate y-axis upper limit
  if (is.null(ymax)) {
    ymax <- max(plot_data$value, na.rm = TRUE)
    if (!is.null(y_break_interval)) {
      yupper <- ceiling(ymax / y_break_interval) * y_break_interval
    } else {
      yupper <- ymax
    }
  } else {
    yupper <- ymax  # Use the provided max
  }
  
  # Create the plot
  p <- ggplot(plot_data, aes(x = category, y = value)) +
    geom_point(color = color, size = 9) +
    labs(
      title = title,
      x = NULL,
      y = NULL
    ) +
    scale_y_continuous(
      limits = c(0, yupper),
      breaks = seq(0, yupper, by = y_break_interval),
      labels = if (!is.null(y_axis_labels)) y_axis_labels else waiver(),
      expand = expansion(mult = c(0, 0.05))
    ) +
    theme_classic() +
    theme(
      plot.title = element_text(
        size = title_size,
        face = "bold", 
        vjust = 1,
        hjust = 0.5, 
        margin = margin(t = -20, b = 60)
      ),
      plot.margin = margin(45, 40, 25, 40),
      axis.text.x = element_text(size = x_text, color = "black"),
      axis.text.y = element_text(size = y_text, color = "black"),
      axis.ticks.x = element_blank(),
      axis.ticks.y = element_line(color = "black"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    ) +
    annotation_custom(
      grob = textGrob(
        label = y_label, 
        hjust = 0.5,
        vjust = -0.9,
        gp = gpar(fontsize = 14)
      ),
      ymin = yupper,
      ymax = yupper,
      xmin = 0.5,
      xmax = 0.5
    ) +
    coord_cartesian(clip = "off")
  
  # Build the plot and turn off clipping
  gt <- ggplot_gtable(ggplot_build(p))
  gt$layout$clip[gt$layout$name == "panel"] <- "off"
  
  # Save if output file specified
  if (!is.null(output_file)) {
    ggsave(output_file, plot = gt, device = "png", width = 8, height = 5)
  }
  
  # Draw and return
  grid.draw(gt)
  invisible(gt)
}
```

## COLLEGE/CAREER DATA

```{r}
#| label: college-stimuli

# Define category order
class_order <- c("Freshman", "Sophomore", "Junior", "Senior")

# Weekly Study Hours - 3 charts
create_lollipop_chart(
  data = college_career_data,
  category_col = class_year,
  value_col = weekly_study_hours,
  title = "Average Weekly Study\nHours by Class Year",
  y_label = "Study\nHrs./Week",
  category_order = class_order,
  y_break_interval = 5,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_lollipop_hours.png"
)

create_bar_chart(
  data = college_career_data,
  category_col = class_year,
  value_col = weekly_study_hours,
  title = "Average Weekly Study\nHours by Class Year",
  y_label = "Study\nHrs./Week",
  category_order = class_order,
  y_break_interval = 5,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_bar_hours.png"
)

create_dot_plot(
  data = college_career_data,
  category_col = class_year,
  value_col = weekly_study_hours,
  title = "Average Weekly Study\nHours by Class Year",
  y_label = "Study\nHrs./Week",
  category_order = class_order,
  y_break_interval = 5,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_dot_hours.png"
)

# Grad School Plans - 3 charts
create_lollipop_chart(
  data = college_career_data,
  category_col = class_year,
  value_col = intention_attend_grad_school_pct,
  title = "Percent of Students Planning\nto Attend Graduate School by Class Year",
  y_label = "% Planning to\nAttend Grad School",
  category_order = class_order,
  y_break_interval = 20,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_lollipop_grad.png",
  title_size = 14
)

create_bar_chart(
  data = college_career_data,
  category_col = class_year,
  value_col = intention_attend_grad_school_pct,
  title = "Percent of Students Planning\nto Attend Graduate School by Class Year",
  y_label = "% Planning to\nAttend Grad School",
  category_order = class_order,
  y_break_interval = 20,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_bar_grad.png",
  title_size = 14
)

create_dot_plot(
  data = college_career_data,
  category_col = class_year,
  value_col = intention_attend_grad_school_pct,
  title = "Percent of Students Planning\nto Attend Graduate School by Class Year",
  y_label = "% Planning to\nAttend Grad School",
  category_order = class_order,
  y_break_interval = 20,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_dot_grad.png",
  title_size = 14
)

# Career Center Satisfaction - 3 charts
create_lollipop_chart(
  data = college_career_data,
  category_col = class_year,
  value_col = satisfaction_career_center,
  title = "Average Satisfaction with\nCareer Services by Class Year",
  y_label = "Avg.\nSatisfaction\n(1-5)",
  category_order = class_order,
  ymax = 5,
  y_break_interval = 1,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_lollipop_satisfaction.png",
  y_text = 10
)

create_bar_chart(
  data = college_career_data,
  category_col = class_year,
  value_col = satisfaction_career_center,
  title = "Average Satisfaction with\nCareer Services by Class Year",
  y_label = "Avg. Satisfaction\n(1-5)",
  category_order = class_order,
  ymax = 5,
  y_break_interval = 1,
  y_text = 10,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_bar_satisfaction.png"
)

create_dot_plot(
  data = college_career_data,
  category_col = class_year,
  value_col = satisfaction_career_center,
  title = "Average Satisfaction with\nCareer Services by Class Year",
  y_label = "Avg. Satisfaction\n(1-5)",
  category_order = class_order,
  ymax = 5,
  y_break_interval = 1,
  color = "seagreen",
  output_file = "trisha-stimuli-pngs/college_dot_satisfaction.png",
  y_text = 10
)
```

## GARDEN DATA

```{r}
#| label: garden-stimuli

# Produce Per Volunteer - 3 charts
create_lollipop_chart(
  data = garden_data,
  category_col = site,
  value_col = produce_handled_per_volunteer,
  title = "Average Produce Handled\nper Volunteer by Garden Site",
  y_label = "Avg. Produce\n(kg)/Volunteer",
  y_break_interval = 1,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_lollipop_produce.png"
)

create_bar_chart(
  data = garden_data,
  category_col = site,
  value_col = produce_handled_per_volunteer,
  title = "Average Produce Handled\nper Volunteer by Garden Site",
  y_label = "Avg. Produce\n(kg)/Volunteer",
  y_break_interval = 1,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_bar_produce.png"
)

create_dot_plot(
  data = garden_data,
  category_col = site,
  value_col = produce_handled_per_volunteer,
  title = "Average Produce Handled\nper Volunteer by Garden Site",
  y_label = "Avg. Produce\n(kg)/Volunteer",
  y_break_interval = 1,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_dot_produce.png"
)

# Volunteer Count - 3 charts
create_lollipop_chart(
  data = garden_data,
  category_col = site,
  value_col = current_volunteer_count,
  title = "Volunteer Counts by Garden Site",
  y_label = "Number of\nVolunteers",
  y_break_interval = 25,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_lollipop_volunteers.png"
)

create_bar_chart(
  data = garden_data,
  category_col = site,
  value_col = current_volunteer_count,
  title = "Volunteer Counts by Garden Site",
  y_label = "Number of\nVolunteers",
  y_break_interval = 25,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_bar_volunteers.png"
)

create_dot_plot(
  data = garden_data,
  category_col = site,
  value_col = current_volunteer_count,
  title = "Volunteer Counts by Garden Site",
  y_label = "Number of\nVolunteers",
  y_break_interval = 25,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_dot_volunteers.png"
)

# Unused Space - 3 charts
create_lollipop_chart(
  data = garden_data,
  category_col = site,
  value_col = unused_physical_space_sqft,
  title = "Unused Physical Space by Garden Site",
  y_label = "Unused Space\n(square feet)",
  y_break_interval = 10,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_lollipop_space.png"
)

create_bar_chart(
  data = garden_data,
  category_col = site,
  value_col = unused_physical_space_sqft,
  title = "Unused Physical Space by Garden Site",
  y_label = "Unused Space\n(square feet)",
  y_break_interval = 10,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_bar_space.png"
)

create_dot_plot(
  data = garden_data,
  category_col = site,
  value_col = unused_physical_space_sqft,
  title = "Unused Physical Space by Garden Site",
  y_label = "Unused Space\n(square feet)",
  y_break_interval = 10,
  color = "chocolate3",
  output_file = "trisha-stimuli-pngs/garden_dot_space.png"
)
```

## LIBRARY DATA

```{r}
#| label: library-stimuli

# Average Books Checked Out per Visit - 3 charts
create_lollipop_chart(
  data = library_data,
  category_col = age_range,
  value_col = avg_books_per_visit,
  title = "Average Books Checked Out\nper Visit by Age Group",
  y_label = "Avg. Books/Visit",
  #y_break_interval = 1,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_lollipop_books.png"
)

create_bar_chart(
  data = library_data,
  category_col = age_range,
  value_col = avg_books_per_visit,
  title = "Average Books Checked Out\nper Visit by Age Group",
  y_label = "Avg.\nBooks/Visit",
  y_break_interval = 1,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_bar_books.png"
)

create_dot_plot(
  data = library_data,
  category_col = age_range,
  value_col = avg_books_per_visit,
  title = "Average Books Checked Out\nper Visit by Age Group",
  y_label = "Avg. Books/Visit",
  ##y_break_interval = 1,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_dot_books.png"
)

# Monthly Visits - 3 charts
create_lollipop_chart(
  data = library_data,
  category_col = age_range,
  value_col = monthly_visits,
  title = "Total Monthly Library Visits by Age Group",
  y_label = "Monthly\nVisits",
  y_break_interval = 100,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_lollipop_monthly.png"
)

create_bar_chart(
  data = library_data,
  category_col = age_range,
  value_col = monthly_visits,
  title = "Total Monthly Library Visits by Age Group",
  y_label = "Monthly\nVisits",
  y_break_interval = 100,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_bar_monthly.png"
)

create_dot_plot(
  data = library_data,
  category_col = age_range,
  value_col = monthly_visits,
  title = "Total Monthly Library Visits by Age Group",
  y_label = "Monthly\nVisits",
  y_break_interval = 100,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_dot_monthly.png"
)

# Total Annual Checkouts - 3 charts
create_lollipop_chart(
  data = library_data |> mutate(total_annual_checkouts = total_annual_checkouts/10),
  category_col = age_range,
  value_col = total_annual_checkouts, 
  title = "Total Annual Book Checkouts by Age Group",
  y_label = "Annual\nCheckouts",
  y_break_interval = 400,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_lollipop_annual.png"
)

create_bar_chart(
  data = library_data |> mutate(total_annual_checkouts = total_annual_checkouts/10),
  category_col = age_range,
  value_col = total_annual_checkouts,
  title = "Total Annual Book Checkouts by Age Group",
  y_label = "Annual\nCheckouts",
  y_break_interval = 400,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_bar_annual.png"
)

create_dot_plot(
  data = library_data |> mutate(total_annual_checkouts = total_annual_checkouts/10),
  category_col = age_range,
  value_col = total_annual_checkouts,
  title = "Total Annual Book Checkouts by Age Group",
  y_label = "Annual\nCheckouts",
  y_break_interval = 400,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_dot_annual.png"
)
```



#| label: library-stimuli

# Number of Cardholders - 3 charts
create_lollipop_chart(
  data = library_data,
  category_col = age_range,
  value_col = num_cardholders,
  title = "Number of Library Cardholders by Age Group",
  y_label = "Number of\nCardholders",
  #y_break_interval = 500,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_cardholders_lollipop.png"
)

create_bar_chart(
  data = library_data,
  category_col = age_range,
  value_col = num_cardholders,
  title = "Number of Library Cardholders by Age Group",
  y_label = "Number of\nCardholders",
  #y_break_interval = 500,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_cardholders_bar.png"
)

create_dot_plot(
  data = library_data,
  category_col = age_range,
  value_col = num_cardholders,
  title = "Number of Library Cardholders by Age Group",
  y_label = "Number of\nCardholders",
  #y_break_interval = 500,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_cardholders_dot.png"
)

# Average Books Checked Out - 3 charts
create_lollipop_chart(
  data = library_data,
  category_col = age_range,
  value_col = avg_books_checked_out,
  title = "Average Library Books Checked Out by Age Group",
  y_label = "Average Books\nChecked Out",
  #y_break_interval = 1,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_books_lollipop.png"
)

create_bar_chart(
  data = library_data,
  category_col = age_range,
  value_col = avg_books_checked_out,
  title = "Average Library Books Checked Out by Age Group",
  y_label = "Average Books\nChecked Out",
  #y_break_interval = 1,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_books_bar.png"
)

create_dot_plot(
  data = library_data,
  category_col = age_range,
  value_col = avg_books_checked_out,
  title = "Average Library Books Checked Out by Age Group",
  y_label = "Average Books\nChecked Out",
  #y_break_interval = 1,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_books_dot.png"
)

# Digital Usage Rate - 3 charts
# Multiply by 100 to convert to percentage
# E-books Checked Out - 3 charts
create_lollipop_chart(
  data = library_data,
  category_col = age_range,
  value_col = digital_usage_rate,
  title = "Proportion of Readers Using E-books by Age Group",
  y_label = "Proportion",
  #y_break_interval = 50,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_ebooks_lollipop.png"
)

create_bar_chart(
  data = library_data,
  category_col = age_range,
  value_col = digital_usage_rate,
  title = "Proportion of Readers Using E-books by Age Group",
  y_label = "Proportion",
  #y_break_interval = 50,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_ebooks_bar.png"
)

create_dot_plot(
  data = library_data,
  category_col = age_range,
  value_col = digital_usage_rate,
  title = "Proportion of Readers Using E-books by Age Group",
  y_label = "Proportion",
  #y_break_interval = 50,
  color = "dodgerblue4",
  output_file = "trisha-stimuli-pngs/library_ebooks_dot.png"
)



## STATE FAIR DATA

```{r}
#| label: fair-stimuli

# Spending - 3 charts
create_lollipop_chart(
  data = fair_data,
  category_col = visitor_type,
  value_col = avg_amount_spent,
  title = "Average Dollars Spent\nby Visitor Type",
  y_label = "Avg. $\nAmount Spent",
  y_break_interval = 200,
  y_axis_labels = scales::label_dollar(),
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_lollipop_spending.png"
)

create_bar_chart(
  data = fair_data,
  category_col = visitor_type,
  value_col = avg_amount_spent,
  title = "Average Dollars Spent\nby Visitor Type",
  y_label = "Avg. $\nAmount Spent",
  y_break_interval = 200,
  y_axis_labels = scales::label_dollar(),
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_bar_spending.png"
)

create_dot_plot(
  data = fair_data,
  category_col = visitor_type,
  value_col = avg_amount_spent,
  title = "Average Dollars Spent\nby Visitor Type",
  y_label = "Avg. $\nAmount Spent",
  y_break_interval = 200,
  y_axis_labels = scales::label_dollar(),
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_dot_spending.png"
)

# All-Days Attendance - 3 charts
create_lollipop_chart(
  data = fair_data,
  category_col = visitor_type,
  value_col = pct_attending_all_days,
  title = "Percent Visitors Who Attended All Days\nof the Fair by Visitor Type",
  y_label = "%\nAttending",
  y_break_interval = 20,
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_lollipop_attendance.png"
)

create_bar_chart(
  data = fair_data,
  category_col = visitor_type,
  value_col = pct_attending_all_days,
  title = "Percent Visitors Who Attended All Days\nof the Fair by Visitor Type",
  y_label = "%\nAttending",
  y_break_interval = 20,
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_bar_attendance.png"
)

create_dot_plot(
  data = fair_data,
  category_col = visitor_type,
  value_col = pct_attending_all_days,
  title = "Percent Visitors Who Attended All Days\nof the Fair by Visitor Type",
  y_label = "%\nAttending",
  y_break_interval = 20,
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_dot_attendance.png"
)

# Ticket Price - 3 charts
create_lollipop_chart(
  data = fair_data,
  category_col = visitor_type,
  value_col = ticket_price,
  title = "Ticket Price by Visitor Type",
  y_label = "Ticket\nPrice",
  y_break_interval = 20,
  y_axis_labels = scales::label_dollar(),
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_lollipop_ticket.png"
)

create_bar_chart(
  data = fair_data,
  category_col = visitor_type,
  value_col = ticket_price,
  title = "Ticket Price by Visitor Type",
  y_label = "Ticket\nPrice",
  y_break_interval = 20,
  y_axis_labels = scales::label_dollar(),
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_bar_ticket.png"
)

create_dot_plot(
  data = fair_data,
  category_col = visitor_type,
  value_col = ticket_price,
  title = "Ticket Price by Visitor Type",
  y_label = "Ticket\nPrice",
  y_break_interval = 20,
  y_axis_labels = scales::label_dollar(),
  color = "steelblue",
  output_file = "trisha-stimuli-pngs/fair_dot_ticket.png"
)
```

##CREDIT DATA

```{r}
#| label: credit-stimuli

# Average Account Balance - 3 charts
# Average Account Balance - 3 charts
create_lollipop_chart(
  data = credit_data |> mutate(balance_thousands = avg_account_balance / 1000),
  category_col = category,
  value_col = balance_thousands,
  title = "Average Account Balance by Client Type",
  y_label = "Average Balance\n(Thousands)",
  y_break_interval = 10,
  color = "darkolivegreen",
  output_file = "trisha-stimuli-pngs/credit_lollipop_balance.png"
)

create_bar_chart(
  data = credit_data |> mutate(balance_thousands = avg_account_balance / 1000),
  category_col = category,
  value_col = balance_thousands,
  title = "Average Account Balance by Client Type",
  y_label = "Average Balance\n(Thousands)",
  y_break_interval = 10,
  color = "darkolivegreen",
  output_file = "trisha-stimuli-pngs/credit_bar_balance.png"
)

create_dot_plot(
  data = credit_data |> mutate(balance_thousands = avg_account_balance / 1000),
  category_col = category,
  value_col = balance_thousands,
  title = "Average Account Balance by Client Type",
  y_label = "Average Balance\n(Thousands)",
  y_break_interval = 10,
  color = "darkolivegreen",
  output_file = "trisha-stimuli-pngs/credit_dot_balance.png"
)

# Churn Rate (Annual Rate of Canceled Accounts) - 3 charts
# Multiply by 100 to convert to percentage
create_lollipop_chart(
  data = credit_data |> mutate(churn_rate_pct = churn_rate * 100),
  category_col = category,
  value_col = churn_rate_pct,
  title = "Annual Rate of Canceled Accounts by Client Type",
  y_label = "Rate of Canceled\nAccounts (%)",
  y_break_interval = 5,
  color = "darkolivegreen",
  output_file = "trisha-stimuli-pngs/credit_lollipop_churn.png"
)

create_bar_chart(
  data = credit_data |> mutate(churn_rate_pct = churn_rate * 100),
  category_col = category,
  value_col = churn_rate_pct,
  title = "Annual Rate of Canceled Accounts by Client Type",
  y_label = "Rate of Canceled\nAccounts (%)",
  y_break_interval = 5,
  color = "darkolivegreen",
  output_file = "trisha-stimuli-pngs/credit_bar_churn.png"
)

create_dot_plot(
  data = credit_data |> mutate(churn_rate_pct = churn_rate * 100),
  category_col = category,
  value_col = churn_rate_pct,
  title = "Annual Rate of Canceled Accounts by Client Type",
  y_label = "Rate of Canceled\nAccounts (%)",
  y_break_interval = 5,
  color = "darkolivegreen",
  output_file = "trisha-stimuli-pngs/credit_dot_churn.png"
)

# Number of Accounts - 3 charts (written out in full thousands)
# Number of Accounts - 3 charts (in thousands with larger intervals)
data <- credit_data |> mutate(customers_thousands = num_customers / 1000)
category_col = data$category
value_col = data$customers_thousands
title = "Number of Accounts by Client Type"
y_label = "Number of\nAccounts\n(Thousands)"
y_break_interval = 30
color = "darkolivegreen"
output_file = "trisha-stimuli-pngs/credit_lollipop_customers.png"

# Set the parameters
data <- credit_data |> mutate(customers_thousands = num_customers / 1000)
category_col <- "category"
value_col <- "customers_thousands"
title <- "Number of Accounts by Client Type"
y_label <- "Number of\nAccounts"
color <- "darkolivegreen"
output_file <- "trisha-stimuli-pngs/credit_lollipop_customers.png"
category_order <- NULL
ymin <- NULL
ymax <- NULL
y_break_interval <- 30
x_text <- 14
y_text <- 14
title_size <- 16

# Prepare the data
plot_data <- data |>
  select(category = category_col, value = value_col)

# Set category order if provided
if (!is.null(category_order)) {
  plot_data <- plot_data |>
    mutate(category = factor(category, levels = category_order))
}

# Calculate y-axis upper limit
if (is.null(ymax)) {
  data_ymax <- max(plot_data$value, na.rm = TRUE)
  if (!is.null(y_break_interval)) {
    yupper <- ceiling(data_ymax / y_break_interval) * y_break_interval
  } else {
    yupper <- data_ymax
  }
} else {
  yupper <- ymax  # Use the provided max
}

# Create the plot
p <- ggplot(plot_data, aes(x = category, y = value)) +
  geom_segment(aes(xend = category, yend = 0),
               color = color, linewidth = 2.3) +
  geom_point(color = color, size = 9) +
  labs(
    title = title,
    x = NULL,
    y = NULL
  ) +
  scale_y_continuous(
    limits = c(0, yupper),
    breaks = seq(0, yupper, by = y_break_interval),
    labels = function(x) ifelse(x == 0, "0", paste0(x, "k")),
    expand = expansion(mult = c(0, 0.05))
  ) +
  
) +
  theme_classic() +
  theme(
    plot.title = element_text(
      size = title_size,
      face = "bold", 
      vjust = 1,
      hjust = 0.5, 
      margin = margin(t = -20, b = 60)
    ),
    plot.margin = margin(45, 40, 25, 40),
    axis.text.x = element_text(size = x_text, color = "black"),
    axis.text.y = element_text(size = y_text, color = "black"),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_line(color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  annotation_custom(
    grob = textGrob(
      label = y_label, 
      hjust = 0.5,
      vjust = -0.9,
      gp = gpar(fontsize = 14)
    ),
    ymin = yupper,
    ymax = yupper,
    xmin = 0.5,
    xmax = 0.5
  ) +
  coord_cartesian(clip = "off")

# Build the plot and turn off clipping
gt <- ggplot_gtable(ggplot_build(p))
gt$layout$clip[gt$layout$name == "panel"] <- "off"

# Save if output file specified
if (!is.null(output_file)) {
  ggsave(output_file, plot = gt, device = "png", width = 8, height = 5)
}

# Draw and return
grid.draw(gt)
invisible(gt)

create_bar_chart(
  data = credit_data |> mutate(customers_thousands = num_customers / 1000),
  category_col = category,
  value_col = customers_thousands,
  title = "Number of Accounts by Client Type",
  y_label = "Number of\nAccounts\n(Thousands)",
  #y_break_interval = 20,
  color = "darkolivegreen",
  output_file = "trisha-stimuli-pngs/credit_bar_customers.png"
)

output_file <- "trisha-stimuli-pngs/credit_bar_customers.png"

# Prepare the data
plot_data <- data |>
  select(category = category_col, value = value_col)

# Set category order if provided
if (!is.null(category_order)) {
  plot_data <- plot_data |>
    mutate(category = factor(category, levels = category_order))
}

# Calculate y-axis upper limit
if (is.null(ymax)) {
  data_ymax <- max(plot_data$value, na.rm = TRUE)
  if (!is.null(y_break_interval)) {
    yupper <- ceiling(data_ymax / y_break_interval) * y_break_interval
  } else {
    yupper <- data_ymax
  }
} else {
  yupper <- ymax  # Use the provided max
}

# Create the plot
p <- ggplot(plot_data, aes(x = category, y = value)) +
  geom_col(fill = color, width = 0.4) +
  labs(
    title = title,
    x = NULL,
    y = NULL
  ) +
  scale_y_continuous(
    limits = c(0, yupper),
    breaks = seq(0, yupper, by = y_break_interval),
    labels = function(x) ifelse(x == 0, "0", paste0(x, "k")),
    expand = expansion(mult = c(0, 0.05))
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(
      size = title_size,
      face = "bold", 
      vjust = 1,
      hjust = 0.5, 
      margin = margin(t = -20, b = 60)
    ),
    plot.margin = margin(45, 40, 25, 40),
    axis.text.x = element_text(size = x_text, color = "black"),
    axis.text.y = element_text(size = y_text, color = "black"),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_line(color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  annotation_custom(
    grob = textGrob(
      label = y_label, 
      hjust = 0.5,
      vjust = -0.9,
      gp = gpar(fontsize = 14)
    ),
    ymin = yupper,
    ymax = yupper,
    xmin = 0.5,
    xmax = 0.5
  ) +
  coord_cartesian(clip = "off")

# Build the plot and turn off clipping
gt <- ggplot_gtable(ggplot_build(p))
gt$layout$clip[gt$layout$name == "panel"] <- "off"

# Save if output file specified
if (!is.null(output_file)) {
  ggsave(output_file, plot = gt, device = "png", width = 8, height = 5)
}

# Draw and return
grid.draw(gt)
invisible(gt)

create_dot_plot(
  data = credit_data |> mutate(customers_thousands = num_customers / 1000),
  category_col = category,
  value_col = customers_thousands,
  title = "Number of Accounts by Client Type",
  y_label = "Number of\nAccounts\n(Thousands)",
  #y_break_interval = 20,
  color = "darkolivegreen",
  output_file = "trisha-stimuli-pngs/credit_dot_customers.png"
)
```


## SCHOOL SUPPLIES DATA

```{r}
#| label: school-stimuli

# Average Price - 3 charts
create_lollipop_chart(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = avg_price,
  title = "Average Price for Each Notebook Type",
  y_label = "Average Price\n(Dollars)",
  y_break_interval = 2,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_lollipop_price.png"
)

create_bar_chart(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = avg_price,
  title = "Average Price for Each Notebook Type",
  y_label = "Average Price\n(Dollars)",
  y_break_interval = 2,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_bar_price.png"
)

create_dot_plot(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = avg_price,
  title = "Average Price for Each Notebook Type",
  y_label = "Average Price\n(Dollars)",
  y_break_interval = 2,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_dot_price.png"
)

# Monthly Sales - 3 charts
create_lollipop_chart(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = monthly_sales,
  title = "Monthly Sales by Notebook Type",
  y_label = "Monthly Sales\n(Units)",
  y_break_interval = 200,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_lollipop_sales.png"
)

create_bar_chart(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = monthly_sales,
  title = "Monthly Sales by Notebook Type",
  y_label = "Monthly Sales\n(Units)",
  y_break_interval = 200,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_bar_sales.png"
)

create_dot_plot(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = monthly_sales,
  title = "Monthly Sales by Notebook Type",
  y_label = "Monthly Sales\n(Units)",
  y_break_interval = 200,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_dot_sales.png"
)

# Return Rate - 3 charts
create_lollipop_chart(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = return_rate_pct,
  title = "Percent of Purchases\nReturned by Notebook Type",
  y_label = "% of Purchases\nReturned",
  y_break_interval = 3,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_lollipop_returns.png"
)

create_bar_chart(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = return_rate_pct,
  title = "Percent of Purchases\nReturned by Notebook Type",
  y_label = "% of Purchases\nReturned",
  y_break_interval = 3,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_bar_returns.png"
)

create_dot_plot(
  data = school_supplies_data,
  category_col = notebook_type,
  value_col = return_rate_pct,
  title = "Percent of Purchases\nReturned by Notebook Type",
  y_label = "% of Purchases\nReturned",
  y_break_interval = 3,
  color = "maroon",
  output_file = "trisha-stimuli-pngs/school_dot_returns.png"
)
```

